% -*- latex -*-
%
%

\subsection{Ganglia}
\label{app:ganglia}

The following is an overview about Ganglia which is taken from
the official website at \url{http://www.ganglia.info}:

``Ganglia is a scalable distributed monitoring system for high-performance
computing systems such as clusters and Grids.  It is based on a
hierarchical design targeted at federations of clusters.  It
leverages widely used technologies such as XML for data representation,
XDR for compact, portable data transport, and RRDtool for data storage
and visualization.  It uses carefully engineered data structures and
algorithms to achieve very low per-node overheads and high concurrency.
The implementation is robust, has been ported to an extensive set of
operating systems and processor architectures, and is currently in use
on over 500 clusters around the world.  It has been used to link
clusters across university campuses and around the world and can scale
to handle clusters with 2000 nodes.''

Ganglia is an included package in OSCAR and it is highly recommended
that you select it to be installed as it is a great tool for monitoring
and visualizing usage of your cluster.

By default, it is assumed that \file{eth0} is the network interface your
compute nodes use to communicate with the headnode.  If this is not
the case, then you will need to configure the package prior to
installation and choose the correct interface.  Other than that,
minimal configuration is required as the other configurable options
are mostly to do with the names of the cluster, grid and owner.
