% -*- latex -*-
%
% Copyright (c) 2002 The Trustees of Indiana University.  
%                    All rights reserved.
% 
% This file is part of the OSCAR software package.  For license
% information, see the COPYING file in the top level directory of the
% OSCAR source distribution.
%
% $Id: download.tex,v 1.5 2002/11/16 19:24:16 jsquyres Exp $
%
% $COPYRIGHT$
%

\section{Downloading an OSCAR Distribution Package}
\label{sec:download}

The OSCAR distribution packages can be downloaded from:

\vspace{11pt}
\centerline{\url{http://oscar.sourceforge.net/}}
\vspace{11pt}

Note that there are actually three flavors of distribution packages
for OSCAR, depending on your bandwidth and installation/development
needs:

\begin{enumerate}
\item ``Regular'': All the OSCAR installation material that most users
  need to install and operate an OSCAR cluster.
  
\item ``Extra Crispy'': Same as Regular, except that the SRPMs for
  most of the RPMs in OSCAR are also included.  SRPMs are {\em not}
  required for installation or normal operation of an OSCAR cluster.
  This distribution package is significantly larger than Regular, and
  is not necessary for most users -- only those who are interested in
  source RPMs need the Extra Crispy distribution.  The SRPMs can be
  found under \file{packages/*/SRPMS/} directories.
  
\item ``Secret Sauce'': This distribution contains {\em only} the
  SRPMs for RPMs in OSCAR in the Regular and Extra Crispy
  distributions (it's essentially (Extra Crispy - Regular)).  It is
  intended only for those who initially downloaded the Regular
  distribution and later decided that they wanted the SRPMs as well.
  The Secret Sauce distribution is intended to be expanded over your
  Regular installation -- it will create \file{packages/*/SRPMS/}
  directories and populate them with the relevant \file{.src.rpm}
  files.

\item ``Low Fat'': This distribution contains {\em only} the OSCAR core. The
OSCAR core is composed of packages providing tools for cluster installation and
management. All the packages providing user applications are removed.
\end{enumerate}

All four distributions can be downloaded from the main OSCAR web
page.


% LocalWords:  Exp
