% -*- latex -*-
%
% Copyright (c) 2002 The Trustees of Indiana University.  
%                    All rights reserved.
% 
% This file is part of the OSCAR software package.  For license
% information, see the COPYING file in the top level directory of the
% OSCAR source distribution.
%
% $Id: intro.tex,v 1.18 2002/11/16 19:24:16 jsquyres Exp $
%
% $COPYRIGHT$
%

\section{Introduction}

The OSCAR (Open Source Cluster Application Resource) software package
is intended to simplify the complex tasks required to install a
cluster.  While the usual intended use for OSCAR clusters is for
high-performance computing (HPC), OSCAR clusters can be used for any
cluster-enabled kinds of applications.  Note that since OSCAR is aimed
towards HPC, several HPC-related packages are installed by default,
such as popular MPI implementations, PVM, PBS, etc.

This document provides a step-by-step installation guide for system
administrators, as well as a detailed explanation of what is happening
as you install.  Note that this installation guide is specific to
OSCAR version \oscarversion.  

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Latest Documentation}

Please be sure that you have the latest version of this document.  It
is possible (and probable!) that newer versions of this document were
released on the main OSCAR web site after the software was
released. You are {\em strongly} encouraged to check
\url{http://oscar.sourceforge.net} for the latest version of these
instructions before proceeding.  Document versions can be compared by
checking their version number and date on the cover page.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Terminology}

A common term used in this document is \term{cluster}, which refers to
a group of individual computers bundled together using hardware and
software in order to make them work as a single machine.

Each individual machine of a cluster is referred to as a \term{node}.
Within the OSCAR cluster to be installed, there are two types of
nodes: \term{server} and \term{client}. A \term{server} node is
responsible for servicing the requests of \term{client} nodes.  A
\term{client} node is dedicated to computation.  

An OSCAR cluster consists of one server node and one or more client
nodes, where all the client nodes [currently] must have homogeneous
hardware.  The software contained within OSCAR does support doing
multiple cluster installs from the same server, but no documentation
is provided on how to do so.

An \term{OSCAR package} is a set of files that is used to install a
software package in an OSCAR cluster.  An OSCAR package can be as
simple as a single RPM file, or it can be more complex, perhaps
including a mixture of RPM and other auxilliary configuration /
installation files.  OSCAR packages provide the majority of
functionality in OSCAR clusters.  

OSCAR packages fall into one of three categories:

\begin{itemize}
\item \term{Core packages} are required for the operation of OSCAR
  itself (mostly involved with the installer).
  
\item \term{Included packages} are shipped in the official OSCAR
  distribution.  These are usually authored and/or packaged by OSCAR
  developers, and have some degreee of official testing before
  release.
  
\item \term{Third party packages} are not included in the official
  OSCAR distribution; they are ``add-ons'' that can be unpacked in the
  OSCAR tree, and therefore installed using the OSCAR installation
  framework.  
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Supported Distributions}

OSCAR has been tested to work with several distributions.
Table~\ref{tab:oscar-distro-support} lists each distribution and
version and specifies the level of support for each. In order to
ensure a successful installation, most users should stick to a
distribution that is listed as \emph{Fully supported}.

\input{supported}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Minimum System Requirements}
\label{sec:intro-min-sys}

\begchange

The following is a list of minimum system requirements for the OSCAR
server node:

\begin{itemize}
\item CPU of i586 or above
\item A network interface card that supports a TCP/IP stack
\item If your OSCAR server node is going to be the router between a
  public network and the cluster nodes, you will need a second
  network interface card that supports a TCP/IP stack
\item At least 4GB total free space -- 2GB under \file{/} and 2GB
  under \file{/var}
\item An installed version of Linux, preferably a {\em Fully
    supported} distribution from Table~\ref{tab:oscar-distro-support}
\end{itemize}

\noindent The following is a list of minimum system requirements for
the OSCAR client nodes:

\begin{itemize}
\item CPU of i586 or above
\item A disk on each client node, at least 2GB in size (OSCAR will
  format the disks during the installation)
\item A network interface card that supports a TCP/IP
  stack\footnote{Beware of certain models of 3COM cards -- not all
    models of 3COM cards are supported by the installation Linux
    kernel that is shipped with OSCAR.  See the OSCAR web site for
    more information. \label{foot:3com-warning}}
\item Same Linux distribution and version as the server node
\item All clients must have the same architecture (e.g., ia32 vs.\ 
  ia64)
\item Monitors and keyboards may be helpful, but are not required
\end{itemize}

\endchange

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Document Organization}

There are both ``Quick Start'' and detailed installation sections in
this guide. Due to the complicated nature of putting together a
high-performance cluster, it is strongly suggested that even
experienced administrators read this document through, without
skipping any sections, and then use the detailed installation
procedure to install your OSCAR cluster.  Novice users will be
comforted to know that anyone who has installed and used Linux can
successfully navigate through the OSCAR cluster install.  If you are
too impatient to read manuals, and your cluster meets certain
requirements, feel free to try your luck with the ``Quick Start''
installation.

The rest of this document is organized as follows.  
%
First, Section~\ref{sec:download} tells how to obtain an OSCAR version
\oscarversion\ distribution package.
%
Next, the ``Release Notes'' section (Section~\ref{sec:release-notes})
that applies to OSCAR version \oscarversion\ contains some
requirements and update issues that need to be resolved before the
install.
%
Section~\ref{sec:sis} provides an overview for the System Installation
Suite software package used in OSCAR to perform the bulk of the
cluster installation.
%
Section~\ref{sec:detail} details the cluster installation procedure
(the level of detail lies somewhere between ``the install will now
update some files'' and ``the install will now replace the string
`xyz' with `abc' in file \file{some\_file}.'')
%
Finally, Section~\ref{sec:pkg-specific-notes} contains system
administration notes about several of the individual packages that are
installed by OSCAR.  {\bf\em This section is a ``must read'' for all
  OSCAR system administrators.}

Appendix~\ref{app:net-boot-client-nodes} covers the topic of network
booting client nodes, which is so important that it deserved its own
section.  
%
Appendix~\ref{app:client-install} provides curious users an overview
of what really happens during a client install.
%
Appendix~\ref{app:no_dhcp} discusses how to install an OSCAR cluster
without a DHCP server.
%
Appendix~\ref{app:security} covers a primer of some security aspects
of a Linux cluster.  Although not intended to be a comprehensive
description of cluster security, it is a good overview for those who
know relatively little about system administration and security.
%
Finally, Appendix~\ref{app:screen-by-screen} is a screen-by-screen
walk through of a typical OSCAR installation.

More information is available on the OSCAR web site and archives of
the various OSCAR mailing lists.  If you have a question that cannot
be answered by this document (including answers to common installation
problems), be sure to visit the OSCAR web site:

\vspace{11pt}
\centerline{\url{http://oscar.sourceforge.net/}}

% LocalWords:  Exp
